---
title: "Project Model"
author: "Lynn Fan"
date: "4/26/2020"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(tidyverse)
library(dplyr)
library(glmnet)
library(car)
library(MASS)
library(data.table)
library(bayesm)
library(ggplot2)
library(R2admb)
library(glmmADMB)
library(lme4)
library(tidyr)
library(mcmc)
library(dplyr)
library(reshape2)
library(bayesplot)
library(varhandle)
library(loo)
```

## Data Cleanup
```{r data cleanup}
data <- read.table("rawdata.txt", 
               col.names=c('stops', 'pop', 'past.arrests', 'precinct', 'eth', 'crime'), 
               fill=FALSE, 
               strip.white=TRUE)
```


## Exploratory Data Analysis
```{r dispersion}
r <- c(mean(data$stops), var(data$stops))
c(mean=r[1], var=r[2], ratio=r[2]/r[1])
```
Overdispersed, so we should do Negative Binomial instead of Poisson.

```{r raneffect}
png('ran_effect.png')
data %>% 
  group_by(precinct) %>%
  ggplot(., mapping = aes(x = as.factor(precinct), y = stops)) +
    geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust=1)) +
  labs(title="number of stops by precincts", x="precincts")
dev.off()
```

```{r regression}
stops<-data$stops ; ethi<-as.factor(data$eth) ; precinct<-as.factor(data$precinct);arrest=data$past.arrests
overdisp_fun <- function(model) {
    rdf <- df.residual(model)
    rp <- residuals(model,type="pearson")
    Pearson.chisq <- sum(rp^2)
    prat <- Pearson.chisq/rdf
    pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
    c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}
# Poisson with random effects
fit.poi <- glmer(stops~1+ethi+(1|precinct),family = poisson(link = "log"), nAGQ = 100)
summary(fit.poi)
overdisp_fun(fit.poi)
# Negative Binomial
fit.nb <- glmer.nb(stops~1+ethi+(1|precinct), verbose=TRUE)
summary(fit.nb)
overdisp_fun(fit.nb)
```

```{r hierarchial}
n <- nrow(data)
precinct.number <- unique(data$precinct)
n.precinct <- length(precinct.number)
precincts <- rep(NA,n)
pblack <- rep(NA,n.precinct)
for (i in 1:n.precinct) {
  temp <- data[data$precinct==i,]
  blackpop <- temp[temp$eth==1,]$pop[1]
  totalpop <- temp[temp$eth==1,]$pop[1]+temp[temp$eth==2,]$pop[1]+temp[temp$eth==3,]$pop[1]
  pblack[i]<-blackpop/totalpop
}
precinct.category <- ifelse (pblack < .1, 1, ifelse (pblack < .4, 2, 3))
arrests <- data$past.arrests
dcjs <- log(arrests*15/12)
dcjs[which(!is.finite(dcjs))] <- 0
crime <- data$crime
pop <- data$pop
stop_df <- as.data.frame (cbind(stops, ethi, precinct, crime, precinct.category, arrests, dcjs))
stop_df$ethi <- as.factor(ethi)
# Multilevel analysis of NYC police stops

# lmer() fits
M1 <- as.list (rep (NA, 12))
M2 <- as.list (rep (NA, 12))
index <- 0
for (j in 1:3){
  for (k in 1:4){
    index <- index + 1
    ok <- precinct.category==j & crime==k
    M1[[index]] <- glmer(stops~1+dcjs+ethi+(1|precinct), #Poisson with random effect
     family=poisson(link="log"), subset=ok, data=stop_df)
    #Negative Binomial
    M2[[index]] <- glmer.nb(stops~1+dcjs+ethi+(1|precinct), verbose=TRUE,subset=ok,data=stop_df,nAGQ=0)
  }
}
M1[1]
M2[1]
anova(M1[[2]],M2[[2]])
```
negative binomial with overdispersion effect 

how do we know proposal distribution? MVN
Poisson GLM with random effects
If NB: link function for negative binomial, try a bayesian glm package, see the parametization, and change the link
  for negative binomial, the r parameter:
  you need to sample both from beta and r for posterior sampling
  
prior for beta: MVN or whatever in the package
prior for r: uninformative uniform distribution

```{r stop-cleanup}
stop_clean <- as.data.frame(cbind(stop_df$stops,stop_df$precinct.category,stop_df$crime,stop_df$dcjs,stop_df$arrests,to.dummy(stop_df$ethi, "ethi")))
colnames(stop_clean) <- c("stops","precinct.category","crime","dcjs","arrests","black","hispanic","white")
``` 

```{r bayesm-1, echo=FALSE}
R=1000
set.seed(66)
simnegbin = function(X, beta, alpha) {
  # Simulate from the Negative Binomial Regression
  lambda = exp(X%*%beta)
  y = NULL
  for (j in 1:length(lambda)) { y = c(y, rnbinom(1, mu=lambda[j], size=alpha)) }
  return(y)
}
data1<-stop_clean[precinct.category==1 & crime==1,]
nobs = nrow(data1)
nvar = 3 # Number of X variables
alpha = 5
Vbeta = diag(nvar)*0.01
# Construct the regdata (containing X)
simnegbindata = data1
beta = c(-1.1052,1.0529, -0.2125,-0.7315) #what we get from negative binomial regression
X<-cbind(rep(1,length(data1$stops)),data1$dcjs,data1$hispanic,data1$white)
simnegbindata = list(y=simnegbin(X,beta,alpha), X=X, beta=beta)
Data1 = simnegbindata
Mcmc1 = list(R=R)
out = rnegbinRw(Data=Data1, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out$alphadraw, tvalues=alpha)
summary(out$betadraw, tvalues=beta)
## plotting examples
out$acceptrbeta
```

```{r predictive check}
betadraws <- out$betadraw #posterior beta
alphadraws <- out$alphadraw #posterior alpha
z.mcmc <- NULL
# posterior predictive
for(i in 1:nrow(betadraws)){ 
  z <- simnegbin(X,betadraws[i,],alphadraws[i]) #sampling from the posterior
  z.mcmc <- rbind(z.mcmc, z)
}
ppc_dens_overlay(data1$stops, z.mcmc[940:1000,]) 
ppc_stat_2d(data1$stops, z.mcmc, stat = c("mean", "var"))

#### Figure 10.5 (traceplots & ACF)
par(mar=c(2.75,2.75,.5,.5),mgp=c(1.7,.7,0))
par(mfrow=c(1,3))
blabs<-c(expression(betadraws[,1]),expression(betadraws[,2]),expression(betadraws[,3],expression(betadraws[,4])))
thin<-c(1,(1:100)*(R/100))
j<-4
plot(thin,betadraws[thin,j],type="l",xlab="iteration",ylab=blabs[j])
abline(h=mean(betadraws[,j]) )

acf(betadraws[,j],ci.col="gray",xlab="lag")
acf(betadraws[thin,j],xlab="lag/10",ci.col="gray") #ACF of thinned chain 
```

```{r bayesian-2, echo=FALSE}
data2<-stop_clean[precinct.category==2 & crime==1,]
nobs = nrow(data2)
# Construct the regdata (containing X)
simnegbindata2 = data2
beta2 = c(-0.2951,0.9412,-0.3686,-1.0353)
X2<-cbind(rep(1,length(data2$stops)),data2$dcjs,data2$hispanic,data2$white)
simnegbindata2 = list(y=simnegbin(X2,beta2,alpha), X=X2, beta=beta2)
Data2 = simnegbindata2
Mcmc2 = list(R=R)
out2 = rnegbinRw(Data=Data2, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out2$alphadraw, tvalues=alpha)
summary(out2$betadraw, tvalues=beta2)

betadraws2 <- out$betadraw
alphadraws2 <- out$alphadraw
z.mcmc2 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws2)){ 
  z <- simnegbin(X2,betadraws2[i,],alphadraws2[i])
  z.mcmc2 <- rbind(z.mcmc2, z)
}
ppc_dens_overlay(data2$stops, z.mcmc2[940:1000,]) 
ppc_stat_2d(data2$stops, z.mcmc2, stat = c("mean", "var"))

#### Figure 10.5 (traceplots & ACF)
par(mar=c(2.75,2.75,.5,.5),mgp=c(1.7,.7,0))
par(mfrow=c(1,3))
blabs2<-c(expression(betadraws2[,1]),expression(betadraws2[,2]),expression(betadraws2[,3],expression(betadraws2[,4])))
thin<-c(1,(1:100)*(R/100))
j<-4
plot(thin,betadraws2[thin,j],type="l",xlab="iteration",ylab=blabs2[j])
abline(h=mean(betadraws2[,j]) )

acf(betadraws2[,j],ci.col="gray",xlab="lag")
acf(betadraws2[thin,j],xlab="lag/10",ci.col="gray") #ACF of thinned chain 
```

```{r bayesian-3, echo=FALSE}
data3<-stop_clean[precinct.category==3 & crime==1,]
nobs = nrow(data3)
# Construct the regdata (containing X)
simnegbindata3 = data3
beta3 <- c(0.3637,0.8602,-0.5615,-1.0791)
X3<-cbind(rep(1,length(data3$stops)),data3$dcjs,data3$hispanic,data3$white)
simnegbindata3 = list(y=simnegbin(X3,beta3,alpha), X=X3, beta=beta3)
Data3 = simnegbindata3
Mcmc3 = list(R=R)
out3 = rnegbinRw(Data=Data3, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out3$alphadraw, tvalues=alpha)
summary(out3$betadraw, tvalues=beta3)

betadraws3 <- out3$betadraw
alphadraws3 <- out3$alphadraw
z.mcmc3 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws3)){ 
  z <- simnegbin(X3,betadraws3[i,],alphadraws3[i])
  z.mcmc3 <- rbind(z.mcmc3, z)
}
ppc_dens_overlay(data3$stops, z.mcmc3[940:1000,]) 
ppc_stat_2d(data3$stops, z.mcmc3, stat = c("mean", "var"))

#### Figure 10.5 (traceplots & ACF)
par(mar=c(2.75,2.75,.5,.5),mgp=c(1.7,.7,0))
par(mfrow=c(1,3))
blabs3<-c(expression(betadraws3[,1]),expression(betadraws3[,2]),expression(betadraws3[,3],expression(betadraws3[,4])))
thin<-c(1,(1:100)*(R/100))
j<-4
plot(thin,betadraws3[thin,j],type="l",xlab="iteration",ylab=blabs3[j])
abline(h=mean(betadraws3[,j]) )

acf(betadraws3[,j],ci.col="gray",xlab="lag")
acf(betadraws3[thin,j],xlab="lag/10",ci.col="gray") #ACF of thinned chain 
```

```{r bayesian-4, echo=FALSE}
data4<-stop_clean[precinct.category==1 & crime==2,]
nobs = nrow(data4)
# Construct the regdata (containing X)
simnegbindata4 = data4
beta4 <- c(0.3158, 1.0483 ,-0.1337,-0.7454)
X4<-cbind(rep(1,length(data4$stops)),data4$dcjs,data4$hispanic,data4$white)
simnegbindata4 = list(y=simnegbin(X4,beta4,alpha), X=X4, beta=beta4)
Data4 = simnegbindata4
Mcmc4 = list(R=R)
out4 = rnegbinRw(Data=Data4, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out4$alphadraw, tvalues=alpha)
summary(out4$betadraw, tvalues=beta4)

betadraws4 <- out4$betadraw
alphadraws4 <- out4$alphadraw
z.mcmc4 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws4)){ 
  z <- simnegbin(X4,betadraws4[i,],alphadraws4[i])
  z.mcmc4 <- rbind(z.mcmc4, z)
}
ppc_dens_overlay(data4$stops, z.mcmc4[940:1000,]) 
ppc_stat_2d(data4$stops, z.mcmc4, stat = c("mean", "var"))

#### Figure 10.5 (traceplots & ACF)
par(mar=c(2.75,2.75,.5,.5),mgp=c(1.7,.7,0))
par(mfrow=c(1,3))
blabs4<-c(expression(betadraws4[,1]),expression(betadraws4[,2]),expression(betadraws4[,3],expression(betadraws4[,4])))
thin<-c(1,(1:100)*(R/100))
j<-4
plot(thin,betadraws4[thin,j],type="l",xlab="iteration",ylab=blabs4[j])
abline(h=mean(betadraws4[,j]) )

acf(betadraws4[,j],ci.col="gray",xlab="lag")
acf(betadraws4[thin,j],xlab="lag/10",ci.col="gray") #ACF of thinned chain 
```


```{r bayesian-5, echo=FALSE}
data5<-stop_clean[precinct.category==1 & crime==3,]
nobs = nrow(data5)
# Construct the regdata (containing X)
simnegbindata5 = data5
beta5 <- c(-0.849341,0.961556,0.166298, 0.004212)
X5<-cbind(rep(1,length(data5$stops)),data5$dcjs,data5$hispanic,data5$white)
simnegbindata5 = list(y=simnegbin(X5,beta5,alpha), X=X5, beta=beta5)
Data5 = simnegbindata5
Mcmc5 = list(R=R)
out5 = rnegbinRw(Data=Data5, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out5$alphadraw, tvalues=alpha)
summary(out5$betadraw, tvalues=beta5)

betadraws5 <- out5$betadraw
alphadraws5 <- out5$alphadraw
z.mcmc5 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws5)){ 
  z <- simnegbin(X5,betadraws5[i,],alphadraws5[i])
  z.mcmc5 <- rbind(z.mcmc5, z)
}
ppc_dens_overlay(data5$stops, z.mcmc5[940:1000,]) 
ppc_stat_2d(data5$stops, z.mcmc5, stat = c("mean", "var"))

#### Figure 10.5 (traceplots & ACF)
par(mar=c(2.75,2.75,.5,.5),mgp=c(1.7,.7,0))
par(mfrow=c(1,3))
blabs5<-c(expression(betadraws5[,1]),expression(betadraws5[,2]),expression(betadraws5[,3],expression(betadraws5[,4])))
thin<-c(1,(1:100)*(R/100))
j<-4
plot(thin,betadraws5[thin,j],type="l",xlab="iteration",ylab=blabs5[j])
abline(h=mean(betadraws5[,j]) )

acf(betadraws5[,j],ci.col="gray",xlab="lag")
acf(betadraws5[thin,j],xlab="lag/10",ci.col="gray") #ACF of thinned chain 
```




```{r bayesian-6, echo=FALSE}
data6<-stop_clean[precinct.category==1 & crime==4,]
nobs = nrow(data6)
# Construct the regdata (containing X)
simnegbindata6 = data6
beta6 <- c(0.0846, 0.6786,0.1546,-0.3825)
X6<-cbind(rep(1,length(data6$stops)),data6$dcjs,data6$hispanic,data6$white)
simnegbindata6 = list(y=simnegbin(X6,beta6,alpha), X=X6, beta=beta6)
Data6 = simnegbindata6
Mcmc6 = list(R=R)
out6 = rnegbinRw(Data=Data6, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out6$alphadraw, tvalues=alpha)
summary(out6$betadraw, tvalues=beta6)

betadraws6 <- out6$betadraw
alphadraws6 <- out6$alphadraw
z.mcmc6 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws6)){ 
  z <- simnegbin(X6,betadraws6[i,],alphadraws6[i])
  z.mcmc6 <- rbind(z.mcmc5, z)
}
ppc_dens_overlay(data6$stops, z.mcmc6[940:1000,]) 
ppc_stat_2d(data6$stops, z.mcmc6, stat = c("mean", "var"))

#### Figure 10.5 (traceplots & ACF)
par(mar=c(2.75,2.75,.5,.5),mgp=c(1.7,.7,0))
par(mfrow=c(1,3))
blabs6<-c(expression(betadraws6[,1]),expression(betadraws6[,2]),expression(betadraws6[,3],expression(betadraws6[,4])))
thin<-c(1,(1:100)*(R/100))
j<-4
plot(thin,betadraws6[thin,j],type="l",xlab="iteration",ylab=blabs6[j])
abline(h=mean(betadraws6[,j]) )

acf(betadraws6[,j],ci.col="gray",xlab="lag")
acf(betadraws6[thin,j],xlab="lag/10",ci.col="gray") #ACF of thinned chain 
```

```{r bayesian-7, echo=FALSE}
data7<-stop_clean[precinct.category==2 & crime==2,]
nobs = nrow(data7)
# Construct the regdata (containing X)
simnegbindata7 = data7
beta7 <- c(0.8441,0.9329,-8.705e-05,-0.5229) #M2[[6]]
X7<-cbind(rep(1,length(data7$stops)),data7$dcjs,data7$hispanic,data7$white)
simnegbindata7 = list(y=simnegbin(X7,beta7,alpha), X=X7, beta=beta7)
Data7 = simnegbindata7
Mcmc7 = list(R=R)
out7 = rnegbinRw(Data=Data7, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out7$alphadraw, tvalues=alpha)
summary(out7$betadraw, tvalues=beta7)

betadraws7 <- out7$betadraw
alphadraws7 <- out7$alphadraw
z.mcmc7 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws7)){ 
  z <- simnegbin(X7,betadraws7[i,],alphadraws7[i])
  z.mcmc7 <- rbind(z.mcmc7, z)
}
ppc_dens_overlay(data7$stops, z.mcmc7[940:1000,]) 
ppc_stat_2d(data7$stops, z.mcmc7, stat = c("mean", "var"))

#### Figure 10.5 (traceplots & ACF)
par(mar=c(2.75,2.75,.5,.5),mgp=c(1.7,.7,0))
par(mfrow=c(1,3))
blabs7<-c(expression(betadraws7[,1]),expression(betadraws7[,2]),expression(betadraws7[,3],expression(betadraws7[,4])))
thin<-c(1,(1:100)*(R/100))
j<-4
plot(thin,betadraws7[thin,j],type="l",xlab="iteration",ylab=blabs7[j])
abline(h=mean(betadraws7[,j]) )

acf(betadraws7[,j],ci.col="gray",xlab="lag")
acf(betadraws7[thin,j],xlab="lag/10",ci.col="gray") #ACF of thinned chain 
```


```{r bayesian-8, echo=FALSE}
data8<-stop_clean[precinct.category==2 & crime==3,]
nobs = nrow(data8)
# Construct the regdata (containing X)
simnegbindata8 = data8
beta8 <- c(0.2109,0.7563,0.4820,0.2032 ) #M2[[7]]
X8<-cbind(rep(1,length(data8$stops)),data8$dcjs,data8$hispanic,data8$white)
simnegbindata8 = list(y=simnegbin(X8,beta8,alpha), X=X8, beta=beta8)
Data8 = simnegbindata8
Mcmc8 = list(R=R)
out8 = rnegbinRw(Data=Data8, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out8$alphadraw, tvalues=alpha)
summary(out8$betadraw, tvalues=beta8)

betadraws8 <- out8$betadraw
alphadraws8 <- out8$alphadraw
z.mcmc8 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws8)){ 
  z <- simnegbin(X8,betadraws8[i,],alphadraws8[i])
  z.mcmc8 <- rbind(z.mcmc8, z)
}
ppc_dens_overlay(data8$stops, z.mcmc8[940:1000,]) 
ppc_stat_2d(data8$stops, z.mcmc8, stat = c("mean", "var"))

#### Figure 10.5 (traceplots & ACF)
par(mar=c(2.75,2.75,.5,.5),mgp=c(1.7,.7,0))
par(mfrow=c(1,3))
blabs8<-c(expression(betadraws8[,1]),expression(betadraws8[,2]),expression(betadraws8[,3],expression(betadraws8[,4])))
thin<-c(1,(1:100)*(R/100))
j<-4
plot(thin,betadraws8[thin,j],type="l",xlab="iteration",ylab=blabs8[j])
abline(h=mean(betadraws8[,j]) )

acf(betadraws8[,j],ci.col="gray",xlab="lag")
acf(betadraws8[thin,j],xlab="lag/10",ci.col="gray") #ACF of thinned chain 
```

```{r bayesian-9, echo=FALSE}
data9<-stop_clean[precinct.category==2 & crime==4,]
nobs = nrow(data9)
# Construct the regdata (containing X)
simnegbindata9 = data9
beta9 <- c(-0.26012,0.72009,0.09347,-0.13026) #M2[[8]]
X9<-cbind(rep(1,length(data9$stops)),data9$dcjs,data9$hispanic,data9$white)
simnegbindata9 = list(y=simnegbin(X9,beta9,alpha), X=X9, beta=beta9)
Data9 = simnegbindata9
Mcmc9 = list(R=R)
out9 = rnegbinRw(Data=Data9, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out9$alphadraw, tvalues=alpha)
summary(out9$betadraw, tvalues=beta9)

betadraws9 <- out9$betadraw
alphadraws9 <- out9$alphadraw
z.mcmc9 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws9)){ 
  z <- simnegbin(X9,betadraws9[i,],alphadraws9[i])
  z.mcmc9 <- rbind(z.mcmc9, z)
}
ppc_dens_overlay(data9$stops, z.mcmc9[940:1000,]) 
ppc_stat_2d(data9$stops, z.mcmc9, stat = c("mean", "var"))

#### Figure 10.5 (traceplots & ACF)
par(mar=c(2.75,2.75,.5,.5),mgp=c(1.7,.7,0))
par(mfrow=c(1,3))
blabs9<-c(expression(betadraws9[,1]),expression(betadraws9[,2]),expression(betadraws9[,3],expression(betadraws9[,4])))
thin<-c(1,(1:100)*(R/100))
j<-4
plot(thin,betadraws9[thin,j],type="l",xlab="iteration",ylab=blabs9[j])
abline(h=mean(betadraws9[,j]) )

acf(betadraws9[,j],ci.col="gray",xlab="lag")
acf(betadraws9[thin,j],xlab="lag/10",ci.col="gray") #ACF of thinned chain 
```

```{r bayesian-10, echo=FALSE}
data10<-stop_clean[precinct.category==3 & crime==2,]
nobs = nrow(data10)
# Construct the regdata (containing X)
simnegbindata10 = data10
beta10 <- c(0.98258,0.93954,-0.09601,-0.60514 ) #M2[[10]]
X10<-cbind(rep(1,length(data10$stops)),data10$dcjs,data10$hispanic,data10$white)
simnegbindata10 = list(y=simnegbin(X10,beta10,alpha), X=X10, beta=beta10)
Data10 = simnegbindata10
Mcmc10 = list(R=R)
out10 = rnegbinRw(Data=Data10, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out10$alphadraw, tvalues=alpha)
summary(out10$betadraw, tvalues=beta10)

betadraws10 <- out10$betadraw
alphadraws10 <- out10$alphadraw
z.mcmc10 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws10)){ 
  z <- simnegbin(X10,betadraws10[i,],alphadraws10[i])
  z.mcmc10 <- rbind(z.mcmc10, z)
}
ppc_dens_overlay(data10$stops, z.mcmc10[940:1000,]) 
ppc_stat_2d(data10$stops, z.mcmc10, stat = c("mean", "var"))

#### Figure 10.5 (traceplots & ACF)
par(mar=c(2.75,2.75,.5,.5),mgp=c(1.7,.7,0))
par(mfrow=c(1,3))
blabs10<-c(expression(betadraws10[,1]),expression(betadraws10[,2]),expression(betadraws10[,3],expression(betadraws10[,4])))
thin<-c(1,(1:100)*(R/100))
j<-4
plot(thin,betadraws10[thin,j],type="l",xlab="iteration",ylab=blabs10[j])
abline(h=mean(betadraws10[,j]) )

acf(betadraws10[,j],ci.col="gray",xlab="lag")
acf(betadraws10[thin,j],xlab="lag/10",ci.col="gray") #ACF of thinned chain 
```

```{r bayesian-11, echo=FALSE}
data11<-stop_clean[precinct.category==3 & crime==3,]
nobs = nrow(data11)
# Construct the regdata (containing X)
simnegbindata11 = data11
beta11 <- c(0.2157,0.7845, 0.1470,0.3764) #M2[[11]]
X11<-cbind(rep(1,length(data11$stops)),data11$dcjs,data11$hispanic,data11$white)
simnegbindata11 = list(y=simnegbin(X11,beta11,alpha), X=X11, beta=beta11)
Data11 = simnegbindata11
Mcmc11 = list(R=R)
out11 = rnegbinRw(Data=Data11, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out11$alphadraw, tvalues=alpha)
summary(out11$betadraw, tvalues=beta11)

betadraws11 <- out11$betadraw
alphadraws11 <- out11$alphadraw
z.mcmc11 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws11)){ 
  z <- simnegbin(X11,betadraws11[i,],alphadraws11[i])
  z.mcmc11 <- rbind(z.mcmc11, z)
}
ppc_dens_overlay(data11$stops, z.mcmc11[940:1000,]) 
ppc_stat_2d(data11$stops, z.mcmc11, stat = c("mean", "var"))

#### Figure 10.5 (traceplots & ACF)
par(mar=c(2.75,2.75,.5,.5),mgp=c(1.7,.7,0))
par(mfrow=c(1,3))
blabs11<-c(expression(betadraws11[,1]),expression(betadraws11[,2]),expression(betadraws11[,3],expression(betadraws11[,4])))
thin<-c(1,(1:100)*(R/100))
j<-4
plot(thin,betadraws11[thin,j],type="l",xlab="iteration",ylab=blabs11[j])
abline(h=mean(betadraws11[,j]) )

acf(betadraws11[,j],ci.col="gray",xlab="lag")
acf(betadraws11[thin,j],xlab="lag/10",ci.col="gray") #ACF of thinned chain 
```

```{r bayesian-12, echo=FALSE}
data12<-stop_clean[precinct.category==3 & crime==4,]
nobs = nrow(data12)
# Construct the regdata (containing X)
simnegbindata12 = data12
beta12 <- c(-0.37151,0.74176,-0.09273,-0.59405) #M2[[12]]
X12<-cbind(rep(1,length(data12$stops)),data12$dcjs,data12$hispanic,data12$white)
simnegbindata12 = list(y=simnegbin(X12,beta12,alpha), X=X12, beta=beta12)
Data12 = simnegbindata12
Mcmc12 = list(R=R)
out12 = rnegbinRw(Data=Data12, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out12$alphadraw, tvalues=alpha)
summary(out12$betadraw, tvalues=beta12)

betadraws12 <- out12$betadraw
alphadraws12 <- out12$alphadraw
z.mcmc12 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws12)){ 
  z <- simnegbin(X12,betadraws12[i,],alphadraws12[i])
  z.mcmc12 <- rbind(z.mcmc12, z)
}
ppc_dens_overlay(data12$stops, z.mcmc12[940:1000,]) 
ppc_stat_2d(data12$stops, z.mcmc12, stat = c("mean", "var"))

#### Figure 10.5 (traceplots & ACF)
par(mar=c(2.75,2.75,.5,.5),mgp=c(1.7,.7,0))
par(mfrow=c(1,3))
blabs12<-c(expression(betadraws12[,1]),expression(betadraws12[,2]),expression(betadraws12[,3],expression(betadraws12[,4])))
thin<-c(1,(1:100)*(R/100))
j<-4
plot(thin,betadraws12[thin,j],type="l",xlab="iteration",ylab=blabs12[j])
abline(h=mean(betadraws12[,j]) )

acf(betadraws12[,j],ci.col="gray",xlab="lag")
acf(betadraws12[thin,j],xlab="lag/10",ci.col="gray") #ACF of thinned chain 
```

