---
title: "Project Model"
author: "Lynn Fan"
date: "4/26/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(tidyverse)
library(dplyr)
library(glmnet)
library(car)
library(MASS)
library(data.table)
library(bayesm)
library(ggplot2)
library(R2admb)
library(glmmADMB)
library(lme4)
library(tidyr)
library(mcmc)
library(dplyr)
library(reshape2)
library(bayesplot)
library(varhandle)
library(loo)
```

## Data Cleanup
```{r data cleanup}
data <- read.table("rawdata.txt", 
               col.names=c('stops', 'pop', 'past.arrests', 'precinct', 'eth', 'crime'), 
               fill=FALSE, 
               strip.white=TRUE)
```


## Exploratory Data Analysis
```{r dispersion}
r <- c(mean(data$stops), var(data$stops))
c(mean=r[1], var=r[2], ratio=r[2]/r[1])
```
Overdispersed, so we should do Negative Binomial instead of Poisson.

```{r regression}
stops<-data$stops ; ethi<-as.factor(data$eth) ; precinct<-as.factor(data$precinct);arrest=data$past.arrests
overdisp_fun <- function(model) {
    rdf <- df.residual(model)
    rp <- residuals(model,type="pearson")
    Pearson.chisq <- sum(rp^2)
    prat <- Pearson.chisq/rdf
    pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
    c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}
# Poisson with random effects
fit.poi <- glmer(stops~1+ethi+(1|precinct),family = poisson(link = "log"), nAGQ = 100)
summary(fit.poi)
overdisp_fun(fit.poi)
# Negative Binomial
fit.nb <- glmer.nb(stops~1+ethi+(1|precinct), verbose=TRUE)
summary(fit.nb)
overdisp_fun(fit.nb)
```

```{r hierarchial}
n <- nrow(data)
precinct.number <- unique(data$precinct)
n.precinct <- length(precinct.number)
precincts <- rep(NA,n)
pblack <- rep(NA,n.precinct)
for (i in 1:n.precinct) {
  temp <- data[data$precinct==i,]
  blackpop <- temp[temp$eth==1,]$pop[1]
  totalpop <- temp[temp$eth==1,]$pop[1]+temp[temp$eth==2,]$pop[1]+temp[temp$eth==3,]$pop[1]
  pblack[i]<-blackpop/totalpop
}
precinct.category <- ifelse (pblack < .1, 1, ifelse (pblack < .4, 2, 3))
arrests <- data$past.arrests
dcjs <- log(arrests*15/12)
dcjs[which(!is.finite(dcjs))] <- 0
crime <- data$crime
pop <- data$pop
stop_df <- as.data.frame (cbind(stops, ethi, precinct, crime, precinct.category, arrests, dcjs))
stop_df$ethi <- as.factor(ethi)
# Multilevel analysis of NYC police stops

# lmer() fits
M1 <- as.list (rep (NA, 12))
M2 <- as.list (rep (NA, 12))
index <- 0
for (j in 1:3){
  for (k in 1:4){
    index <- index + 1
    ok <- precinct.category==j & crime==k
    M1[[index]] <- glmer(stops~1+dcjs+ethi+(1|precinct), #Poisson with random effect
     family=poisson(link="log"), subset=ok, data=stop_df)
    #Negative Binomial
    M2[[index]] <- glmer.nb(stops~1+dcjs+ethi+(1|precinct), verbose=TRUE,subset=ok,data=stop_df,nAGQ=0)
  }
}
M1[1]
M2[1]
anova(M1[[2]],M2[[2]])
```
negative binomial with overdispersion effect 

how do we know proposal distribution? MVN
Poisson GLM with random effects
If NB: link function for negative binomial, try a bayesian glm package, see the parametization, and change the link
  for negative binomial, the r parameter:
  you need to sample both from beta and r for posterior sampling
  
prior for beta: MVN or whatever in the package
prior for r: uninformative uniform distribution

```{r stop-cleanup}
stop_clean <- as.data.frame(cbind(stop_df$stops,stop_df$precinct.category,stop_df$crime,stop_df$dcjs,stop_df$arrests,to.dummy(stop_df$ethi, "ethi")))
colnames(stop_clean) <- c("stops","precinct.category","crime","dcjs","arrests","black","hispanic","white")
``` 

```{r bayesm-1}
R=1000
set.seed(66)
simnegbin = function(X, beta, alpha) {
  # Simulate from the Negative Binomial Regression
  lambda = exp(X%*%beta)
  y = NULL
  for (j in 1:length(lambda)) { y = c(y, rnbinom(1, mu=lambda[j], size=alpha)) }
  return(y)
}
data1<-stop_clean[precinct.category==1 & crime==1,]
nobs = nrow(data1)
nvar = 3 # Number of X variables
alpha = 5
Vbeta = diag(nvar)*0.01
# Construct the regdata (containing X)
simnegbindata = data1
beta = c(-1.1052,1.0529, -0.2125,-0.7315) #what we get from negative binomial regression
X<-cbind(rep(1,length(data1$stops)),data1$dcjs,data1$hispanic,data1$white)
simnegbindata = list(y=simnegbin(X,beta,alpha), X=X, beta=beta)
Data1 = simnegbindata
Mcmc1 = list(R=R)
out = rnegbinRw(Data=Data1, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out$alphadraw, tvalues=alpha)
summary(out$betadraw, tvalues=beta)
## plotting examples
out$acceptrbeta
```

```{r predictive check}
betadraws <- out$betadraw #posterior beta
alphadraws <- out$alphadraw #posterior alpha
z.mcmc <- NULL
# posterior predictive
for(i in 1:nrow(betadraws)){ 
  z <- simnegbin(X,betadraws[i,],alphadraws[i]) #sampling from the posterior
  z.mcmc <- rbind(z.mcmc, z)
}
ppc_dens_overlay(data1$stops, z.mcmc[940:1000,]) 
ppc_stat_2d(data1$stops, z.mcmc, stat = c("mean", "var"))
```

```{r bayesian-2}
data2<-stop_clean[precinct.category==2 & crime==1,]
nobs = nrow(data2)
# Construct the regdata (containing X)
simnegbindata2 = data2
beta2 = c(-0.2951,0.9412,-0.3686,-1.0353)
X2<-cbind(rep(1,length(data2$stops)),data2$dcjs,data2$hispanic,data2$white)
simnegbindata2 = list(y=simnegbin(X2,beta2,alpha), X=X2, beta=beta2)
Data2 = simnegbindata2
Mcmc2 = list(R=R)
out2 = rnegbinRw(Data=Data2, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out2$alphadraw, tvalues=alpha)
summary(out2$betadraw, tvalues=beta2)

betadraws2 <- out$betadraw
alphadraws2 <- out$alphadraw
z.mcmc2 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws2)){ 
  z <- simnegbin(X2,betadraws2[i,],alphadraws2[i])
  z.mcmc2 <- rbind(z.mcmc2, z)
}
ppc_dens_overlay(data2$stops, z.mcmc2[940:1000,]) 
ppc_stat_2d(data2$stops, z.mcmc2, stat = c("mean", "var"))
```

```{r bayesian-3}
data3<-stop_clean[precinct.category==3 & crime==1,]
nobs = nrow(data3)
# Construct the regdata (containing X)
simnegbindata3 = data3
beta3 <- c(0.3637,0.8602,-0.5615,-1.0791)
X3<-cbind(rep(1,length(data3$stops)),data3$dcjs,data3$hispanic,data3$white)
simnegbindata3 = list(y=simnegbin(X3,beta3,alpha), X=X3, beta=beta3)
Data3 = simnegbindata3
Mcmc3 = list(R=R)
out3 = rnegbinRw(Data=Data3, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out3$alphadraw, tvalues=alpha)
summary(out3$betadraw, tvalues=beta3)

betadraws3 <- out3$betadraw
alphadraws3 <- out3$alphadraw
z.mcmc3 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws3)){ 
  z <- simnegbin(X3,betadraws3[i,],alphadraws3[i])
  z.mcmc3 <- rbind(z.mcmc3, z)
}
ppc_dens_overlay(data3$stops, z.mcmc3[940:1000,]) 
ppc_stat_2d(data3$stops, z.mcmc3, stat = c("mean", "var"))
```

```{r bayesian-4}
data4<-stop_clean[precinct.category==1 & crime==2,]
nobs = nrow(data4)
# Construct the regdata (containing X)
simnegbindata4 = data4
beta4 <- c(0.3158, 1.0483 ,-0.1337,-0.7454)
X4<-cbind(rep(1,length(data4$stops)),data4$dcjs,data4$hispanic,data4$white)
simnegbindata4 = list(y=simnegbin(X4,beta4,alpha), X=X4, beta=beta4)
Data4 = simnegbindata4
Mcmc4 = list(R=R)
out4 = rnegbinRw(Data=Data4, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out4$alphadraw, tvalues=alpha)
summary(out4$betadraw, tvalues=beta4)

betadraws4 <- out4$betadraw
alphadraws4 <- out4$alphadraw
z.mcmc4 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws4)){ 
  z <- simnegbin(X4,betadraws4[i,],alphadraws4[i])
  z.mcmc4 <- rbind(z.mcmc4, z)
}
ppc_dens_overlay(data4$stops, z.mcmc4[940:1000,]) 
ppc_stat_2d(data4$stops, z.mcmc4, stat = c("mean", "var"))
```


```{r bayesian-5}
data5<-stop_clean[precinct.category==1 & crime==3,]
nobs = nrow(data5)
# Construct the regdata (containing X)
simnegbindata5 = data5
beta5 <- c(-0.849341,0.961556,0.166298, 0.004212)
X5<-cbind(rep(1,length(data5$stops)),data5$dcjs,data5$hispanic,data5$white)
simnegbindata5 = list(y=simnegbin(X5,beta5,alpha), X=X5, beta=beta5)
Data5 = simnegbindata5
Mcmc5 = list(R=R)
out5 = rnegbinRw(Data=Data5, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out5$alphadraw, tvalues=alpha)
summary(out5$betadraw, tvalues=beta5)

betadraws5 <- out5$betadraw
alphadraws5 <- out5$alphadraw
z.mcmc5 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws5)){ 
  z <- simnegbin(X5,betadraws5[i,],alphadraws5[i])
  z.mcmc5 <- rbind(z.mcmc5, z)
}
ppc_dens_overlay(data5$stops, z.mcmc4[940:1000,]) 
ppc_stat_2d(data5$stops, z.mcmc5, stat = c("mean", "var"))
```




```{r bayesian-6}
data6<-stop_clean[precinct.category==1 & crime==4,]
nobs = nrow(data6)
# Construct the regdata (containing X)
simnegbindata6 = data6
beta6 <- c(0.0846, 0.6786,0.1546,-0.3825)
X6<-cbind(rep(1,length(data6$stops)),data6$dcjs,data6$hispanic,data6$white)
simnegbindata6 = list(y=simnegbin(X6,beta6,alpha), X=X6, beta=beta6)
Data6 = simnegbindata6
Mcmc6 = list(R=R)
out6 = rnegbinRw(Data=Data6, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out6$alphadraw, tvalues=alpha)
summary(out6$betadraw, tvalues=beta6)

betadraws6 <- out6$betadraw
alphadraws6 <- out6$alphadraw
z.mcmc6 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws6)){ 
  z <- simnegbin(X6,betadraws6[i,],alphadraws6[i])
  z.mcmc6 <- rbind(z.mcmc5, z)
}
ppc_dens_overlay(data6$stops, z.mcmc6[940:1000,]) 
ppc_stat_2d(data6$stops, z.mcmc6, stat = c("mean", "var"))
```

```{r bayesian-7}
data7<-stop_clean[precinct.category==2 & crime==2,]
nobs = nrow(data7)
# Construct the regdata (containing X)
simnegbindata7 = data7
beta7 <- c(0.8441,0.9329,-8.705e-05,-0.5229) #M2[[6]]
X7<-cbind(rep(1,length(data7$stops)),data7$dcjs,data7$hispanic,data7$white)
simnegbindata7 = list(y=simnegbin(X7,beta7,alpha), X=X7, beta=beta7)
Data7 = simnegbindata7
Mcmc7 = list(R=R)
out7 = rnegbinRw(Data=Data7, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out7$alphadraw, tvalues=alpha)
summary(out7$betadraw, tvalues=beta7)

betadraws7 <- out7$betadraw
alphadraws7 <- out7$alphadraw
z.mcmc7 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws7)){ 
  z <- simnegbin(X7,betadraws7[i,],alphadraws7[i])
  z.mcmc7 <- rbind(z.mcmc7, z)
}
ppc_dens_overlay(data7$stops, z.mcmc7[940:1000,]) 
ppc_stat_2d(data7$stops, z.mcmc7, stat = c("mean", "var"))
```


```{r bayesian-8}
data8<-stop_clean[precinct.category==2 & crime==3,]
nobs = nrow(data8)
# Construct the regdata (containing X)
simnegbindata8 = data8
beta8 <- c(0.2109,0.7563,0.4820,0.2032 ) #M2[[7]]
X8<-cbind(rep(1,length(data8$stops)),data8$dcjs,data8$hispanic,data8$white)
simnegbindata8 = list(y=simnegbin(X8,beta8,alpha), X=X8, beta=beta8)
Data8 = simnegbindata8
Mcmc8 = list(R=R)
out8 = rnegbinRw(Data=Data8, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out8$alphadraw, tvalues=alpha)
summary(out8$betadraw, tvalues=beta8)

betadraws8 <- out8$betadraw
alphadraws8 <- out8$alphadraw
z.mcmc8 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws8)){ 
  z <- simnegbin(X8,betadraws8[i,],alphadraws8[i])
  z.mcmc8 <- rbind(z.mcmc8, z)
}
ppc_dens_overlay(data8$stops, z.mcmc8[940:1000,]) 
ppc_stat_2d(data8$stops, z.mcmc8, stat = c("mean", "var"))
```

```{r bayesian-9}
data9<-stop_clean[precinct.category==2 & crime==4,]
nobs = nrow(data9)
# Construct the regdata (containing X)
simnegbindata9 = data9
beta9 <- c(-0.26012,0.72009,0.09347,-0.13026) #M2[[8]]
X9<-cbind(rep(1,length(data9$stops)),data9$dcjs,data9$hispanic,data9$white)
simnegbindata9 = list(y=simnegbin(X9,beta9,alpha), X=X9, beta=beta9)
Data9 = simnegbindata9
Mcmc9 = list(R=R)
out9 = rnegbinRw(Data=Data9, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out9$alphadraw, tvalues=alpha)
summary(out9$betadraw, tvalues=beta9)

betadraws9 <- out9$betadraw
alphadraws9 <- out9$alphadraw
z.mcmc9 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws9)){ 
  z <- simnegbin(X9,betadraws9[i,],alphadraws9[i])
  z.mcmc9 <- rbind(z.mcmc9, z)
}
ppc_dens_overlay(data9$stops, z.mcmc9[940:1000,]) 
ppc_stat_2d(data9$stops, z.mcmc9, stat = c("mean", "var"))
```

```{r bayesian-10}
data10<-stop_clean[precinct.category==3 & crime==2,]
nobs = nrow(data10)
# Construct the regdata (containing X)
simnegbindata10 = data10
beta10 <- c(0.98258,0.93954,-0.09601,-0.60514 ) #M2[[10]]
X10<-cbind(rep(1,length(data10$stops)),data10$dcjs,data10$hispanic,data10$white)
simnegbindata10 = list(y=simnegbin(X10,beta10,alpha), X=X10, beta=beta10)
Data10 = simnegbindata10
Mcmc10 = list(R=R)
out10 = rnegbinRw(Data=Data10, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out10$alphadraw, tvalues=alpha)
summary(out10$betadraw, tvalues=beta10)

betadraws10 <- out10$betadraw
alphadraws10 <- out10$alphadraw
z.mcmc10 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws10)){ 
  z <- simnegbin(X10,betadraws10[i,],alphadraws10[i])
  z.mcmc10 <- rbind(z.mcmc10, z)
}
ppc_dens_overlay(data10$stops, z.mcmc10[940:1000,]) 
ppc_stat_2d(data10$stops, z.mcmc10, stat = c("mean", "var"))
```

```{r bayesian-11}
data11<-stop_clean[precinct.category==3 & crime==3,]
nobs = nrow(data11)
# Construct the regdata (containing X)
simnegbindata11 = data11
beta11 <- c(0.2157,0.7845, 0.1470,0.3764) #M2[[11]]
X11<-cbind(rep(1,length(data11$stops)),data11$dcjs,data11$hispanic,data11$white)
simnegbindata11 = list(y=simnegbin(X11,beta11,alpha), X=X11, beta=beta11)
Data11 = simnegbindata11
Mcmc11 = list(R=R)
out11 = rnegbinRw(Data=Data11, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out11$alphadraw, tvalues=alpha)
summary(out11$betadraw, tvalues=beta11)

betadraws11 <- out11$betadraw
alphadraws11 <- out11$alphadraw
z.mcmc11 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws11)){ 
  z <- simnegbin(X11,betadraws11[i,],alphadraws11[i])
  z.mcmc11 <- rbind(z.mcmc11, z)
}
ppc_dens_overlay(data11$stops, z.mcmc11[940:1000,]) 
ppc_stat_2d(data11$stops, z.mcmc11, stat = c("mean", "var"))
```

```{r bayesian-12}
data12<-stop_clean[precinct.category==3 & crime==4,]
nobs = nrow(data12)
# Construct the regdata (containing X)
simnegbindata12 = data12
beta12 <- c(-0.37151,0.74176,-0.09273,-0.59405) #M2[[12]]
X12<-cbind(rep(1,length(data12$stops)),data12$dcjs,data12$hispanic,data12$white)
simnegbindata12 = list(y=simnegbin(X12,beta12,alpha), X=X12, beta=beta12)
Data12 = simnegbindata12
Mcmc12 = list(R=R)
out12 = rnegbinRw(Data=Data12, Mcmc=list(R=R))
cat("Summary of alpha/beta draw", fill=TRUE)
summary(out12$alphadraw, tvalues=alpha)
summary(out12$betadraw, tvalues=beta12)

betadraws12 <- out12$betadraw
alphadraws12 <- out12$alphadraw
z.mcmc12 <- NULL
# posterior predictive
for(i in 1:nrow(betadraws12)){ 
  z <- simnegbin(X12,betadraws12[i,],alphadraws12[i])
  z.mcmc12 <- rbind(z.mcmc12, z)
}
ppc_dens_overlay(data12$stops, z.mcmc12[940:1000,]) 
ppc_stat_2d(data12$stops, z.mcmc12, stat = c("mean", "var"))
```

### References:
http://www.stat.columbia.edu/~gelman/arm/examples/police/
http://mc-stan.org/bayesplot/reference/bayesplot-package.html
https://rdrr.io/cran/varhandle/man/to.dummy.html
https://www.lakeheadu.ca/sites/default/files/uploads/77/Burke.pdf
https://cran.r-project.org/web/packages/MHadaptive/MHadaptive.pdf
https://rdrr.io/cran/bayesm/man/rnegbinRw.html
http://pcl.missouri.edu/jeff/node/322
http://georglsm.r-forge.r-project.org/site-projects/pdf/Hastings_within_Gibbs.pdf
https://gist.github.com/gaberoo/4619102
https://stats.stackexchange.com/questions/261552/what-is-a-hierarchical-model-that-can-estimated-via-the-metropolis-hastings-algo
https://stats.stackexchange.com/questions/411321/mcmc-metropolis-hastings-sampler-estimation-of-multiple-parameters
http://probability.ca/jeff/ftpdir/adaptex.pdf
https://kieranrcampbell.github.io/blog/2015/03/31/tutorial-bayesian-nb-regression.html
https://rdrr.io/cran/Bolstad2/man/bivnormMH.html
https://rdrr.io/cran/SamplerCompare/src/R/metropolis.R
http://www.stat.umn.edu/geyer/8701/package/mcmc.Rcheck/mcmc/html/metropolis.html
https://arxiv.org/pdf/1308.0657.pdf
https://sakai.duke.edu/access/content/group/cd23a148-0b53-4d1c-9d1e-c66c74e59129/Labs/Lab06/lab09.html#multiclass-logistic-regression
https://blog.abhranil.net/2014/02/08/r-code-for-multivariate-random-walk-metropolis-hastings-sampling/
https://en.wikipedia.org/wiki/Stop-and-frisk_in_New_York_City
https://stats.stackexchange.com/questions/361501/metropolis-hastings-in-a-bayesian-hierarchical-model
https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#fitting-models-with-overdispersion
https://sakai.duke.edu/access/content/group/cd23a148-0b53-4d1c-9d1e-c66c74e59129/Lectures/songbird.R
http://probability.ca/jeff/ftpdir/adaptex.pdf
http://www2.geog.ucl.ac.uk/~mdisney/teaching/GEOGG121/sivia_skilling/mterop_hastings.pdf

```
# Useful variables
y<-stops ; X<-cbind(rep(1,length(y)),dcjs,ethi)
yX<-cbind(y,X)
colnames(yX)<-c("stops","intercept","dcjs","ethi") 
n<-length(y) ; p<-dim(X)[2]
# Prior parameters
pmn.beta<-rep(0,p) # prior mean for beta
psd.beta<-rep(10,p) # prior sd for beta
# Metropolis settings
var.prop<- var(log(y+1/2))*solve( t(X)%*%X ) #variance for proposal distribution for beta
beta<-rep(0,p) #initial beta
r<-runif(1,0,1)#initial r
S<-10000 # no. of MCMC samples
BETA<-matrix(0,nrow=S,ncol=p) #container
R<-rep(0,S)
ac<-0 # no. of accepts in MCMc
set.seed(1)

## rmvnorm function for proposals
rmvnorm<-function(n,mu,Sigma)
{ # samples from the multivariate normal distribution
  E<-matrix(rnorm(n*length(mu)),n,length(mu))
  t(  t(E%*%chol(Sigma)) +c(mu))
}
## MCMC
for(s in 1:S) {
  
  #proposal: sample a candidate beta
  beta.p<- t(rmvnorm(1, beta, var.prop))
  r.p<- rexp(1,1)
  #evaluate: compute log-acceptance-ratio, then accept/reject
  logr <- sum(dnorm(beta.p,pmn.beta,psd.beta,log=T)) - sum(dnorm(beta,pmn.beta,psd.beta,log=T)) +
    dexp(r.p, 1) - dexp(r,1)
  if( log(runif(1))< logr ) { beta<-beta.p ; r<-r.p;ac<-ac+1 }
  
  BETA[s,]<-beta #store sample
  R[s]<-r
}
cat(ac/S,"\n") #acceptance rate of MCMC
library(coda)
apply(BETA,2,effectiveSize) #ESS
apply(R,1,effectiveSize)
```

